Robust Quantum Optimal Control

### 2020-09-07 paper/paper.tex

%% ABSTRACT
\begin{abstract}
  %% - 1 + 2 topic importance
  %% - 1 what I have done
  %% - few sentences about primary results
  The ability to engineer high fidelity operations on quantum processors in the presence of
  decoherence and systematic errors remains the primary challenge requisite to achieving quantum advantage.
  Quantum optimal control (QOC) techniques have been developed to mitigate deleterious effects
  by manipulating the time-dependent parameters controlling the system. These techniques often
  target a single error channel and may be incompatible with experimental constraints. 
  In this work, we develop techniques to simultaneously mitigate
  longitudinal relaxation and pure dephasing in a framework that can handle arbitrary
  experimental constraints.
  Optimizing a gate set for a fluxonium qubit as an example, we achieve a factor of 4 increase in
  longitudinal relaxation times and outperform
  existing dynamic decoupling methods in mitigating pure dephasing due to flux noise, achieving
  high fidelity gates in the presence of parameter spreads up to $5\%$.
\end{abstract}

@@

%% S1
\section{Introduction}
%% P1 - general field, lots of citation. cite reviews, original qoc.
%% P2 - exhaustively find all papers, what people have done to address the issues
%% P3 - what we do is. maybe add a rant about how
%% QOC is classical control theory but we've ignored that field
%% P4 - outline
(The field) The field of quantum optimal control (QOC) is concerned
with improving the efficiency and accuracy with which quantum systems
are manipulated.
Early QOC techniques were proposed for nuclear magnetic resonance experiments
\cite{khaneja2005optimal}, and applications now include superconducting
qubits \cite{heeres2017implementing,
  leng2019robust, leung2017speedup, xu2020nonadiabatic},
neutral atoms and ion traps \cite{van2016optimal}, nitrogen-vacancy centers in
diamonds \cite{rembold2020introduction}, and Bose-Einstien condensates
\cite{sorensen2018quantum}.
In quantum control theory, as in classical control theory,
optimization is performed
to minimize an objective function. Relevant for the field
of quantum computation is the task of achieving high fidelity quantum gates.
In this application, the objective function includes contributions from the
gate error associated with the state transfer of the system as well as contributions
due to the decoherence of the system, such as the evolution time of the gate.
The decision variables of the optimization problem are the time-dependent control
parameters, particular to the quantum system, that govern its evolution.

@@

Numerical quantum control techniques have developed in parallel
with analytic frameworks.
Most numerical techniques proposed for QOC are indirect shooting
methods where an initial guess for the control trajectory is forward-simulated,
e.g. under the time-dependent-Schroedinger-equation,
and first-order gradients are computed for a functional evaluated
on the state of the evolved system \cite{leung2017speedup,  goerz2019krotov, doria2011optimal,
  abdelhafez2019gradient, machnes2015gradient, leng2019robust}.
Most techniques formulate the
problem as unconstrained or use
projective gradient methods to enforce constraints
\cite{machnes2015gradient}.
The latter approach relies on constructing
bases for constraint spaces--which is not always feasible--and
may hinder convergence. Methods for mitigating decoherence using numerical
techniques have employed master equation dynamics (citation needed),
Monte Carlo style trajectories \cite{abdelhafez2019gradient},
and proposals have been made to simulate multiple trajectories
\cite{reinhold2019controlling, rembold2020introduction}.

(This work) In this work, we employ state of the art
trajectory optimization methods
to formulate the quantum
optimal control problem as a constrained optimization problem.
The trajectory optimization framework allows us to handle arbitrary
constraints on the evolution of the quantum state. Additionally,
we introduce methods for constructing gates that mitigate longitudinal
relaxation and are robust to experimental errors, and therefore
pure dephasing, the dominant barriers to experimentally
realizing practical quantum computing.

@@

%% F1
\begin{figure}[ht]
  \begin{subfigure}{.49\linewidth}
    \includegraphics[width=\linewidth]{assets/f1a.png}
  \end{subfigure}\hfill%
  \begin{subfigure}{.49\linewidth}
    \includegraphics[width=\linewidth]{assets/f1b.png}
  \end{subfigure}
  \begin{subfigure}{\linewidth}
    \includegraphics[width=\linewidth]{assets/f1c.png}
  \end{subfigure}
  \caption{(a) $T_{1}$ optimized gates (orange) and analytic gates (blue). (b) Master equation simulation with $T_{1}$ dissipation
    for all gates, over the course of concatenated gate applications. (c) $T_{1}$ interpolation function used in optimization
    and the corresponding $T_{1}$ times for the average absolute amplitude of each gate.}
\end{figure}

@@

(Results) We achieve a factor of 2.5 decrease in the probability
of longitudinal relaxation from the analytic gates we benchmark
against. We simulate the performance of the gates using the
Lindblad master equation (see Appendix). We repeatedly apply the basis gates
and measure the fidelity of the resulting state as a function
of time as shown in Fig. 1.

%% T4.1
%% TODO: Maybe best for appendix.
%% TOOD: Maybe add some commentary about why we see more speedup
%% for Z/2 and X/2 than Y/2. In particular I think this is because of the
%% long idle times at the flux frustration point for Z/2 and X/2. It
%% might be a more fair comparison to show gates that have better T1s
%% and as good flux noise mitigation.

\begin{table}[ht]
  \begin{tabular}{c | c | c | c}
    & Analytic  & QOC &\\
    Gate & $P_{1}\ (10^{-5})$ & $P_{1}\ (10^{-5})$ & Speedup\\
    \hline
    Z/2 & 5.745 & 1.469 & 3.911\\
    Y/2 & 5.253 & 1.400 & 3.752\\
    X/2 & 16.251 & 6.946 & 2.340\\
  \end{tabular}
  \caption{Probability of longitudinal relaxation for each gate
    evaluated at the gate's duration.}
\end{table}

%% F2
\begin{figure}[ht]
  \begin{subfigure}{.49\linewidth}
    \label{fig:5.1:a}
    \includegraphics[width=\linewidth]{assets/f2a.png}
  \end{subfigure}%
  \begin{subfigure}{.49\linewidth}
    \label{fig:5.1:b}
    \includegraphics[width=\linewidth]{assets/f2b.png}
  \end{subfigure}

  \begin{subfigure}{\linewidth}
    \includegraphics[width=\linewidth]{assets/f2c.png}
    \label{fig:5.1:c}
  \end{subfigure}

@@

%% F3
\begin{figure}[ht]
  \begin{subfigure}{.5\linewidth}
    \includegraphics[width=\linewidth]{assets/f3a.png}
  \end{subfigure}%
  \begin{subfigure}{.5\linewidth}
    \includegraphics[width=\linewidth]{assets/f3c.png}
  \end{subfigure}

@@

  \caption{\centering (a) $X/2$ gates robust to flux offsets constructed using the analytic,
    (2, 4)-point sampling, and (2, 3)-derivative methods at a gate duration
    of $56.8 (ns)$. (b) Mean gate error for 10 samples of 1/$f$ flux noise over 5000
    gate applications. (c) Stochastic 1/$f$ flux noise.}
\end{figure}

>>>>

%% ABSTRACT
\begin{abstract}
  %% - 1 + 2 topic importance
  %% - 1 what I have done
  %% - few sentences about primary results
  The ability to engineer high fidelity operations on quantum processors in the presence of
  systematic errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
  Quantum optimal control (QOC) techniques have proven effective in realizing high fidelity operations,
  but they require exquisite calibration to be performant. In this work we employ robust control techniques
  to achieve high fidelity gates in the presence of parameter spreads up to 5$\%$.
  Additionally, current QOC techniques mitigate decoherence effects due to longitudinal relaxation by integrating
  Lindblad master equation dynamics, which drastically increase the computational complexity
  of the optimization algorithm.
  In this work we achieve a factor of $4$ increase in longitudinal relaxation times
  for the gate set of a fluxonium qubit
  using a computationally efficient metric in optimizaiton.

  %% The ability to engineer high fidelity operations on quantum processors in the presence of
  %% systematic errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
  %% Quantum optimal control (QOC) techniques have been developed to mitigate deleterious effects
  %% by manipulating the time-dependent parameters controlling the system. These techniques often
  %% target a single error channel and may be incompatible with experimental constraints. 
  %% In this work, we develop techniques to simultaneously mitigate
  %% longitudinal relaxation and pure dephasing in a framework that can handle arbitrary
  %% experimental constraints.
  %% Optimizing a gate set for a fluxonium qubit as an example, we achieve a factor of 4 increase in
  %% longitudinal relaxation times and outperform
  %% existing dynamic decoupling methods in mitigating pure dephasing due to flux noise, achieving
  %% high fidelity gates in the presence of parameter spreads up to $5\%$.
\end{abstract}

@@

%% S1
\section{Introduction}
textwidth: \printinunitsof{in}\prntlen{\textwidth}
linewidth: \printinunitsof{in}\prntlen{\linewidth}
%% P1 - general field, lots of citation. cite reviews, original qoc.
%% P2 - exhaustively find all papers, what people have done to address the issues
%% P3 - what we do is. maybe add a rant about how
%% QOC is classical control theory but we've ignored that field
%% P4 - outline
(The field) The field of quantum optimal control (QOC) is concerned
with improving the efficiency and accuracy with which quantum systems
are manipulated.
Early QOC techniques were proposed for nuclear magnetic resonance experiments
\cite{khaneja2005optimal}, and applications now include superconducting
qubits \cite{heeres2017implementing,
  leng2019robust, leung2017speedup, xu2020nonadiabatic},
neutral and ionized atoms \cite{van2016optimal}, nitrogen-vacancy centers in
diamond \cite{rembold2020introduction}, and Bose-Einstien condensates
\cite{sorensen2018quantum}.
In quantum control theory, as in classical control theory,
optimization is performed
to minimize an objective function. Relevant for the field
of quantum computation is the task of achieving high fidelity quantum gates.
In this application, the objective function includes contributions from the
gate error associated with the state transfer of the system as well as contributions
due to experimental constraints, such as the control amplitude.
The decision variables of the optimization problem are the time-dependent control
parameters, particular to the quantum system, that govern its evolution.

@@

Numerical quantum control techniques have developed in parallel
with analytic frameworks.
Most numerical techniques proposed for QOC are indirect shooting
methods where an initial guess for the control trajectory is forward-simulated,
e.g. under the time-dependent Schroedinger equation,
and first-order gradients are computed for a functional evaluated
on the state of the evolved system \cite{leung2017speedup,  goerz2019krotov, doria2011optimal,
  abdelhafez2019gradient, machnes2015gradient, leng2019robust}.
Most techniques formulate the
problem as unconstrained or use
projective gradient methods to enforce constraints
\cite{machnes2015gradient}.
The latter approach relies on constructing
bases for constraint spaces--which is not always feasible--and
may hinder convergence. Methods for mitigating decoherence using numerical
techniques have employed master equation dynamics (citation needed),
Monte Carlo style trajectories \cite{abdelhafez2019gradient},
and proposals have been made to simulate multiple trajectories
\cite{reinhold2019controlling, rembold2020introduction}.

(This work) In this work, we employ state of the art
trajectory optimization methods
to formulate the quantum
optimal control problem as a constrained optimization problem, allowing
us to enforce experimental constraints.
Additionally,
we introduce methods for constructing gates that mitigate longitudinal
relaxation and are robust to experimental errors, and therefore
pure dephasing, the dominant barriers to experimentally
realizing practical quantum computing.

@@

%% F1
\begin{figure}[ht]
  \begin{subfigure}{.6\linewidth}
    \includegraphics[width=\linewidth]{assets/f1a.png}
  \end{subfigure}%
    \begin{subfigure}{.4\linewidth}
    \includegraphics[width=\linewidth]{assets/f1b.png}
  \end{subfigure}

  \begin{subfigure}{\linewidth}
    \includegraphics[width=\linewidth]{assets/f1c.png}
  \end{subfigure}
  \caption{
    (a) $T_{1}$ optimized gates (red) and analytic gates (blue).
    (b) $T_{1}$ interpolation function used in optimization
    and the corresponding $T_{1}$ times for the average absolute amplitude of each gate.
    (c) Master equation simulation with $T_{1}$ dissipation
    for all gates, over the course of concatenated gate applications.
  }
\end{figure}

@@

(Results) We achieve a factor of 5 decrease in the probability
of longitudinal relaxation from the analytic gates we benchmark
against. We simulate the performance of the gates using the
Lindblad master equation (see Appendix). We repeatedly apply the basis gates
and measure the fidelity of the resulting state as a function
of time as shown in Fig. 1.

%% T1
%% TODO: maybe for appendix.
\begin{table}[ht]
  \begin{tabular}{c | c | c | c}
    & Analytic  & QOC &\\
    Gate & $P_{1}\ (10^{-5})$ & $P_{1}\ (10^{-5})$ & Ratio\\
    \hline
    Z/2 & 5.745 & 1.149 & 5.000\\
    Y/2 & 5.253 & 1.157 & 4.540\\
    X/2 & 16.251 & 6.946 & 2.340\\
  \end{tabular}
  \caption{Probability of longitudinal relaxation for each gate
    evaluated at the gate's duration.}
\end{table}

%% F2
\begin{figure}[ht]
  \begin{subfigure}{.6\linewidth}
    \includegraphics[width=\linewidth]{assets/f2a.png}
  \end{subfigure}%
  \begin{subfigure}{.4\linewidth}
    \includegraphics[width=\linewidth]{assets/f2b.png}
  \end{subfigure}

  \begin{subfigure}{\linewidth}
    \includegraphics[width=\linewidth]{assets/f2c.png}
  \end{subfigure}

@@

%% F3
\begin{figure}[ht]
  \begin{subfigure}{\linewidth}
    \includegraphics[width=\linewidth]{assets/f3a.png}
  \end{subfigure}

@@

  %% \begin{subfigure}{.5\linewidth}
  %%   \includegraphics[width=\linewidth]{assets/f3c.png}
  %% \end{subfigure}

  \caption{\centering
    (a) $X/2$ gates robust to flux offsets constructed using the analytic,
    (2, 4)-point sampling, and (2, 3)-derivative methods at a gate duration
    of $56.8 (ns)$. (b) Mean gate error for 10 samples of 1/$f$ flux noise over 5000
    gate applications.
    %% (c) Stochastic 1/$f$ flux noise.
  }
\end{figure}
